---
title: "Precision for Medicine - Assessment"
output:
  html_document: default
  html_notebook: default
---

```{r}
install.packages('NanoStringNorm')
install.packages('NanoStringNorm')
install.packages('nlme')
```
```{r}
library(nlme)
library(NanoStringNorm)
library(ggplot2)


```



PREPORCESSING DATA

Annotation file is in the .CSV format. distinguish baseline and post-treatment created two separate directories.
Read annotation file stored into "precision_for_medicine_annotation" data frame. Created two separate data framework for baseline and post-treatment. 

Read baseline and post-treatment RCC files into two different baseline.mRNA and post_treatment.mRNA nanostring variable variable. 

```{r}

precision_for_medicine_annotation <- read.csv2("case_study_datascientist_data/case_study_annotations.csv",header = TRUE,sep=",")
baseline <- subset(precision_for_medicine_annotation,visit == "Baseline")
post_treatment <- subset(precision_for_medicine_annotation, visit =="Post-Treatment")
baseline.mRNA <- read.markup.RCC(rcc.path = "case_study_datascientist_data/base_line",rcc.pattern = "*.RCC|*.rcc",exclude = NULL,include = NULL,nprobes = -1)
post_treatment.mRNA <- read.markup.RCC(rcc.path = "case_study_datascientist_data/post_treatment",rcc.pattern = "*.RCC|*.rcc",exclude = NULL,include = NULL,nprobes = -1)
```



Problem 3.1 (quality control)

Normalize mRNA expression data from the NanoString platform using NanoStringNorm
Create separate NanoStringNorm object for baseline.mRNA and post_treatment.mRNA nanostring variable variable. Data is in the matrix format to generate heat maps. 

```{r}
baseline.mRNA.norm <- NanoStringNorm(
  x = baseline.mRNA,
  anno = NA,
  CodeCount = 'geo.mean',
  Background = 'mean',
  SampleContent = 'housekeeping.geo.mean',
  round.values = TRUE,
  take.log = TRUE,
  return.matrix.of.endogenous.probes = TRUE
)


post_treatment.mRNA.norm <- NanoStringNorm(
  x = post_treatment.mRNA,
  anno = NA,
  CodeCount = 'geo.mean',
  Background = 'mean',
  SampleContent = 'housekeeping.geo.mean',
  round.values = TRUE,
  take.log = TRUE,
  return.matrix.of.endogenous.probes = TRUE
)

```


Heatmap showing positive and negative control genes in columns and samples in rows.
Here, samples are divided into the baseline and post-treatment to evaluate the signal level and noise level accordingly. 
we can also use in one heatmap for all samples.

By considering the data required as matrix format to display we have used 'return.matrix.of.endogenous.probes' while normalizing nanoString heat and use Transpose function to get appropriate format.



```{r}
heatmap(t(baseline.mRNA.norm),main = "Heatmap for Basline Samples") 
heatmap(t(post_treatment.mRNA.norm), main = "Heatmap for Post-Treatment Samples")
```


Problem 3.2 (data analysis)

Subset CXCL1 and MCL1 values from Baseline/Post-Treatment sample and stored into the separate data frame. 

```{r}
 data_CXCL1_MCL1 <- rbind(data.frame(sample = baseline$subjectid ,time_point = baseline$visit,CXCL1 = baseline.mRNA.norm["CXCL1",],MCL1 = baseline.mRNA.norm["MCL1",]),data.frame(sample= post_treatment$subjectid,time_point = post_treatment$visit,CXCL1 = post_treatment.mRNA.norm["CXCL1",],MCL1 = post_treatment.mRNA.norm["MCL1",]))
data_CXCL1_MCL1
```

Display difference from Baseline/Post-Treatment sample. (CXCL1 and MCL1 values from Baseline/Post-Treatment sample)
using BOXPLOT

```{r}



ggplot(data = data_CXCL1_MCL1, aes(x=sample, y=CXCL1))  + geom_boxplot(aes(fill=time_point),stat = "boxplot") + geom_point(aes(fill=time_point),stat = "identity") + ggrepel::geom_text_repel(aes(label = CXCL1), color = "black", size = 2.5, segment.color = "grey") +  labs (x= "Sample Baseline/Post-Treatment", y = "CXCL1") + ggtitle("CXCL1 values differnce between Baseline/Post-Treatment sample") + theme_bw()



 
ggplot(data = data_CXCL1_MCL1, aes(x=sample, y=MCL1))  + geom_boxplot(aes(fill=time_point),stat = "boxplot") + geom_point(aes(fill=time_point)) + geom_boxplot(aes(fill=time_point),stat = "boxplot") + ggrepel::geom_text_repel(aes(label = CXCL1), color = "black", size = 2.5, segment.color = "grey") + labs (x= "Sample Baseline/Post-Treatment", y = "MCL1") + ggtitle("MCL1 values differnce between Baseline/Post-Treatment sample")
 
```

```{r}
summary(data_CXCL1_MCL1)
```

EXTRA/OPTIONAL

Display difference from Baseline/Post-Treatment sample. (CXCL1 and MCL1 values from Baseline/Post-Treatment sample)
using BAR CHART

```{r}

 #CXCL1 values differnce between Baseline/Post-Treatment sample 
 data_CXCL1 <- rbind(data.frame(sample = baseline$subjectid ,time_point = baseline$visit,CXCL1 = baseline.mRNA.norm["CXCL1",]),data.frame(sample= post_treatment$subjectid,time_point = post_treatment$visit,CXCL1 = post_treatment.mRNA.norm["CXCL1",]))
 
ggplot(data_CXCL1, aes(x=sample,y = CXCL1,fill = time_point)) + geom_bar(stat="identity",position=position_dodge(),colour="black") + labs (x= "Sample Baseline/Post-Treatment", y = "CXCL1") + ggtitle("CXCL1 values differnce between Baseline/Post-Treatment sample")
 
 #MCL1 values differnce between Baseline/Post-Treatment sample 
 data_MCL1 <- rbind(data.frame(sample = baseline$subjectid ,time_point = baseline$visit,MCL1 = baseline.mRNA.norm["MCL1",]),data.frame(sample= post_treatment$subjectid,time_point = post_treatment$visit,MCL1 = post_treatment.mRNA.norm["MCL1",]))
 
ggplot(data_MCL1, aes(x=sample,y = MCL1,fill = time_point)) + geom_bar(stat="identity",position=position_dodge(),colour="black") + labs (x= "Sample Baseline/Post-Treatment", y = "MCL1") + ggtitle("MCL1 values differnce between Baseline/Post-Treatment sample")
 


```








Problem 3.3 (reporting)

Normalize mRNA expression data from the NanoString platform using NanoStringNorm.

To get all reports using plot NanoStringNorm function, the data required as the NanoStringNorm format to get all reports. we have used 'return.matrix.of.endogenous.probes' as the FALSE parameter while normalizing nanoString.

```{r}
baseline.mRNA.normN <- NanoStringNorm(
  x = baseline.mRNA,
  anno = NA,
  CodeCount = 'geo.mean',
  Background = 'mean',
  SampleContent = 'housekeeping.geo.mean',
  round.values = TRUE,
  take.log = TRUE,
  return.matrix.of.endogenous.probes = FALSE
)


post_treatment.mRNA.normN <- NanoStringNorm(
  x = post_treatment.mRNA,
  anno = NA,
  CodeCount = 'geo.mean',
  Background = 'mean',
  SampleContent = 'housekeeping.geo.mean',
  round.values = TRUE,
  take.log = TRUE,
  return.matrix.of.endogenous.probes = FALSE
)

```

A brief report with all types:
Here,
Used sperate reports for the Baseline Sample Report and Post Treatment Sample Report. We can use the single report as well. I considered the separate report to understand the report correctly.


```{r}
 Plot.NanoStringNorm(x = baseline.mRNA.normN,label.best.guess = TRUE,plot.type = 'all', title = "Bseline Sample Report")
 Plot.NanoStringNorm(x = post_treatment.mRNA.normN,label.best.guess = TRUE,plot.type = 'all', title = "Post Treatment Sample Report")
 

```

Thanks For consideration !!!